Ceph Parallel File System Evaluation Report

### 2013-09-05 abstract.tex

In collaboration with its major developer, Inktank Inc, we evaluated the Ceph
technology for scientific high-performance computing (HPC) environments.  This paper
presents our evaluation methodology, experiments, results and observations from
mostly parallel I/O performance and scalability perspectives.  Our work made
two unique contributions. First, our evaluation is performed under a realistic
setup for a large-scale capability HPC environment like ours using a commercial
high-end storage system.
% (Data Direct Network's SFA10K).
%although typical in large-scale HPC environment, but atypical for average
%commercial off-the-shelf setup.  
Second, our path of investigation, tuning efforts, and findings made direct
contributions to Ceph's development and improved code quality, scalability, and
performance. These changes should also benefit both Ceph and HPC communities at
large. Throughout the evaluation, we observed that Ceph still is an evolving
technology under fast-paced development and showing great promises.

>>>>

We evaluated the Ceph
technology for scientific high-performance computing (HPC) environments.  This paper
presents our evaluation methodology, experiments, results and observations from
mostly parallel I/O performance and scalability perspectives.  Our work made
two unique contributions. First, our evaluation is performed under a realistic
setup for a large-scale capability HPC environment using a commercial
high-end storage system.
% (Data Direct Network's SFA10K).
%although typical in large-scale HPC environment, but atypical for average
%commercial off-the-shelf setup.  
Second, our path of investigation, tuning efforts, and findings made direct
contributions to Ceph's development and improved code quality, scalability, and
performance. These changes should also benefit both Ceph and HPC communities at
large. Throughout the evaluation, we observed that Ceph still is an evolving
technology under fast-paced development and showing great promises.

### 2013-09-05 baseline.tex
changes here and there

\subsection{Block I/O over Native IB}

The lowest layer in the system are the exposed LUNs to the server hosts from
the RAID controllers.  These LUNs are mostly configured as a RAID6 8+2 array (8
data disks and 2 \textit{P} and \textit{Q} disks). In our earlier tests we
observed that a 7.2K RPM SAS disk can perform at ~140 MB/s for 128 kB
sequential I/O requests and a 7.2K RPM SATA disk can do ~36 MB/s. For these
tests, all disk caches were turned off. Therefore, in 8 data disk RAID group
with 1 MB sequential I/O requests (each disks sees 128 kB chunks of the
request) the expected aggregate performance would be ~1.12 GB/s for a SAS RAID
group and ~288 MB/s for a SATA RAID group, if there are no caches along the I/O
path.  With the caches turned on, especially the write-back cache on the DDN
RAID controller, we observe a significant performance improvement. In our
day-to-day HPC operations we run our storage systems with write-back cache
turned on on RAID controllers, if and only if, there is a method of
cache-mirroring between the two active-active RAID controllers to recover from
a single controller failure scenario. With the write-back cache on we observe
~1.4 GB/s per RAID 6 8+2 SAS group and ~955 MB/s per RAID 6 8+2 SATA group, for
1 MB sequential I/O. These two performance numbers will be used as a baseline
in our study and they are presented in Table~\ref{tbl:block-io-baseline}.  For
these tests, the 8+2 RAID 6 groups are exported to the server hosts using SCSI
RDMA Protocol (SRP) over IB protocol (which is the default for our normal HPC
configurations). For a SATA disk group this translates into roughly 120
MB/s/disk and for a SAS disk group it is 175 MB/s. As can be seen, SATA RAID
groups benefit from caching more than SAS groups.

@@

As mentioned earlier, we had 280 SATA disks and 200 SAS disks. Organizing these
into RAID 6 8+2 groups will yield 28 SATA and 20 SAS RAID 6 8+2 groups. We had
4 server hosts in our testbed and evenly distributing these on to 4 server
hosts translates into 8 SATA LUNs and 4 SAS LUNs per server host (when a RAID
group is exported to a server host, it is labeled as a LUN from the server host
point of view).

Next, we established the baseline performance for a single server host. We
exercised all seven SATA LUNs concurrently form a single server host at the
block-level. This resulted in an aggregate performance of roughly 2.6 GB/s.
Repeating the same test using only four SAS LUNs yielded around 2.8 GB/s
aggregate performance for 1 MB sequential I/O. These two numbers will be used
as single server host aggregate baseline performance for this study. These are
also presented in presented in Table~\ref{tbl:block-io-baseline}.

It is worth mentioning that we used our in-house developed synthetic benchmark,
\textit{fair-lio} for all our block-level testing.  fair-lio has better
sequential I/O characteristics compared to commonly used XDD benchmark and it
is asynchronous and based on the Linux libaio library.

\begin{table}[htb]
\centering
\caption{Block I/O performances on single LUN and single host}
\label{tbl:block-io-baseline}

\begin{tabular}{| l | l |}
    \hline
    SAS single LUN sequential read & ~1.4 GB/s \\
    SATA single LUN sequential read & ~955 MB/s \\
    Single host with four SAS LUNs & ~ 2.8 GB/s \\
    Single host with seven SATA LUNs & ~ 2.6 GB/s \\
    \hline
\end{tabular}
\end{table}

When exercising all 28 SATA LUNs from all four server hosts in parallel, we
observed an 11 GB/s aggregate performance. Same level of performance was
observed using all 20 SAS LUNs from four server hosts in parallel. Comparing
these to the advertised peak performance of 12 GB/s of the DDN SFA10K, we
concluded that our test setup is well configured to drive the backend disk
system at full performance and we are limited by the RAID controller
performance. Going forward in this study, 11 GB/s will be used the peak
baseline for the entire test configuration.

>>>>

\subsection{Block I/O over Native IB} 
\label{sec:block-io}

The lowest layer in the system are the exposed LUNs to the server hosts from
the RAID controllers.  These LUNs are configured as a RAID 6 8+2 array (8 data
disks and 2 \textit{P} and \textit{Q} disks). In our earlier tests we observed
that a 7.2K RPM SAS disk can perform at \textasciitilde140 MB/s for 128 kB
sequential I/O requests and a 7.2K RPM SATA disk can do \textasciitilde36 MB/s.
For these tests, all disk caches were turned off. Therefore, in 8 data disk
RAID group with 1 MB sequential I/O requests (each disks sees 128 kB chunks of
the request) the expected aggregate performance would be \textasciitilde1.12
GB/s for a SAS RAID group and \textasciitilde288 MB/s for a SATA RAID group, if
there are no caches along the I/O path.  With the caches turned on, especially
the write-back cache on the DDN RAID controller, we observe a significant
performance improvement. In our day-to-day HPC operations we run our storage
systems with write-back cache turned on on RAID controllers, if and only if,
there is a method of cache-mirroring between the two active-active RAID
controllers. With the write-back cache on we observe \textasciitilde1.4 GB/s
per RAID 6 8+2 SAS group and  \textasciitilde955 MB/s per RAID 6 8+2 SATA
group, for 1 MB sequential I/O.  These two performance numbers will be used as
a baseline in our study and they are presented in
Table~\ref{tbl:block-io-baseline}.  For these tests, the 8+2 RAID 6 groups are
exported to the server hosts using SCSI RDMA Protocol (SRP) over IB protocol
(which is the default for our normal HPC configurations). For a SATA disk group
this translates into roughly 120 MB/s/disk and for a SAS disk group it is 175
MB/s/disk. As can be seen, SATA RAID groups benefit from caching more than SAS
groups.

@@

As mentioned earlier, we had 280 SATA disks and 200 SAS disks. Organizing these
into RAID 6 8+2 groups will yield 28 SATA and 20 SAS RAID 6 8+2 groups. We had
4 server hosts in our testbed and evenly distributing these on to 4 server
hosts translates into 7 SATA LUNs and 4 SAS LUNs per server host.

Next, we established the baseline performance for a single server host. We
exercised all seven SATA LUNs concurrently form a single server host at the
block-level. This resulted in an aggregate performance of roughly 2.6 GB/s.
Repeating the same test using only four SAS LUNs yielded around 2.8 GB/s
aggregate performance for 1 MB sequential I/O. These two numbers will be used
as single server host aggregate baseline performance for this study. These are
also presented in Table~\ref{tbl:block-io-baseline}.

It is worth mentioning that we used our in-house developed synthetic benchmark,
\textit{fair-lio} for all our block-level testing.  fair-lio has better
sequential I/O characteristics compared to commonly used XDD benchmark and it
is asynchronous and based on the Linux \verb!libaio! library.

\begin{table}[htb]
\centering
\caption{Baseline block I/O performance summary}
\label{tbl:block-io-baseline}

\begin{tabular}{ l | l }
    \hline
    SAS single LUN sequential read & ~1.4 GB/s \\
    SATA single LUN sequential read & ~955 MB/s \\
    Single host with four SAS LUNs & ~ 2.8 GB/s \\
    Single host with seven SATA LUNs & ~ 2.6 GB/s \\
    \hline
\end{tabular}
\end{table}

When exercising all 28 SATA LUNs from all four server hosts in parallel, we
observed an 11 GB/s aggregate performance. Same level of performance was
observed using all 20 SAS LUNs from four server hosts in parallel. Comparing
these to the advertised peak performance of 12 GB/s of the DDN SFA10K, we
concluded that our test setup is well configured to drive the backend disk
system at close to full performance and we are limited by the RAID controller
performance. Going forward in this study, 11 GB/s will be used the peak
baseline for the entire test configuration.

### 2013-09-05 introduction.tex

Data objects are distributed across storage daemons (ceph-osd) using
CRUSH\cite{crush}, a deterministic hashing function that allows administrators
to define flexible placement policies over a hierarchical cluster structure
(e.g., disks, hosts, racks, rows, datacenters).  Because the location of
objects can be calculated based on the object identifier and cluster layout
(similar to consistent hashing~\cite{..}), there is no need for a metadata
index or server for the RADOS object store. Further, because CRUSH provides an
authoritative view of data placement, storage daemons can coordinate directly
to handle data replications, recovery, or object migratation in the face of
failure or cluster topology changes.  A small cluster of monitors (ceph-mon
daemons) use Paxos to provide consensus on the current cluster layout, but do
not need to explicitly coordinate migration or recovery activities.

@@

Client hosts that mount the file system communicate with one or more metadata
servers to traverse the namespace but perform file IO by reading and writing
directly to RADOS objects that contain the file data.  The metadata server
cluster monitors the temperature of cached metadata and periodically adjusts
the distribution of the namespace across the MDS cluster by migrating
responsibility for arbitrary subtrees of the hierarchy between a dynamic pool
of active ceph-mds daemons.  This dynamic subtree partitioning~\cite{..}
strategy is both adaptive and highly scalable, allowing additional metadata
server daemons to be added or removed at any time, making it ideally suited
both for large-scale workloads with bursty workloads or general purpose
clusters whose workloads grow or contract over time.

>>>>

Data objects are distributed across storage daemons (ceph-osd) using
CRUSH\cite{Weil:2006:Crush}, a deterministic hashing function that allows
administrators to define flexible placement policies over a hierarchical
cluster structure (e.g., disks, hosts, racks, rows, datacenters).  Because the
location of objects can be calculated based on the object identifier and
cluster layout (similar to consistent hashing~\cite{karger1997consistent}),
there is no need for a metadata index or server for the RADOS object store.
Further, because CRUSH provides an authoritative view of data placement,
storage daemons can coordinate directly to handle data replications, recovery,
or object migration in the face of failure or cluster topology changes.  A
small cluster of monitors (ceph-mon daemons) use Paxos to provide consensus on
the current cluster layout, but do not need to explicitly coordinate migration
or recovery activities.

@@

Client hosts that mount the file system communicate with one or more metadata
servers to traverse the namespace but perform file IO by reading and writing
directly to RADOS objects that contain the file data.  The metadata server
cluster monitors the temperature of cached metadata and periodically adjusts
the distribution of the namespace across the MDS cluster by migrating
responsibility for arbitrary subtrees of the hierarchy between a dynamic pool
of active ceph-mds daemons.  This dynamic subtree
partitioning~\cite{Weil:2004:dynamic} strategy is both adaptive and highly
scalable, allowing additional metadata server daemons to be added or removed
at any time, making it ideally suited both for large-scale workloads with
bursty workloads or general purpose clusters whose workloads grow or contract
over time.

### 2013-09-05 testbed.tex

Our Ceph testbed employs a collection of nodes, including the server hosts
(oss-[1-4]). These nodes and their roles are summarized in
Table~\ref{tbl:ceph-test-nodes}. In the following discussion, we use
``servers'', ``osd servers'', ``server hosts'' interchangeably. We will
emphasize with ``client'' prefix when we want to distinguish it from above.

\begin{table}[!t]
\centering
\caption{Support nodes involved in Ceph testbed}
\label{tbl:ceph-test-nodes}
    \begin{tabular}{ll}
    \hline
    Node & Role \\
    \hline
    tick-mds1 & Ceph monitor node \\
    spoon46 & Ceph MDS node \\
    tick-oss[1-4] & Ceph OSD servers \\
    spoon28-31, spoon37-41 & Ceph client nodes \\
    \hline
    \end{tabular}
\end{table}

All hosts  (client and servers) were configured with Redhat 6.3 and kernel
version 3.5.1 initially, and later upgraded to 3.9, Glibc
2.12 with syncfs support, locally patched.  We used the Ceph 0.48 and 0.55
release in the initial tests, upgraded to 0.64 and then to 0.67RC for a final
round of tests.

@@

Our tools for benchmarking and evaluating different layers also change as we go
up in the stack.  For example, a tool perfect for assessing block-level
performance over a single server host might be completely inadequate for
observing parallel file system performance over multiple clients. As we present
our evaluations, we will also introduce and explain our benchmark tools, in
this paper.

>>>>

\begin{comment}
Our Ceph testbed employs a collection of nodes, including the server hosts
(oss-[1-4]). These nodes and their roles are summarized in
Table~\ref{tbl:ceph-test-nodes}. 
\end{comment}

In the following discussion, we use ``servers'', ``osd servers'', ``server
hosts'' interchangeably. We will emphasize with ``client'' prefix when we want
to distinguish it from above.  All hosts  (client and servers) were configured
with Redhat 6.3 and kernel version 3.5.1 initially, and later upgraded to 3.9,
Glibc 2.12 with syncfs support, locally patched.  We used the Ceph 0.48 and
0.55 release in the initial tests, upgraded to 0.64 and then to 0.67RC for a
final round of tests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% REMOVED
\begin{comment}
\begin{table}[!t]
\centering
\caption{Support nodes involved in Ceph testbed}
\label{tbl:ceph-test-nodes}
    \begin{tabular}{ll}
    \hline
    Node & Role \\
    \hline
    tick-mds1 & Ceph monitor node \\
    spoon46 & Ceph MDS node \\
    tick-oss[1-4] & Ceph OSD servers \\
    spoon28-31, spoon37-41 & Ceph client nodes \\
    \hline
    \end{tabular}
\end{table}
\end{comment}

@@

Our tools for benchmarking and evaluating different layers also change as we
go up in the stack.  For example, a tool perfect for assessing block-level
performance over a single server host might be completely inadequate for
observing parallel file system performance over multiple clients. As we
present our evaluations, we will also introduce and explain our benchmark
tools along the way.
